Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-04-05 articles/formalism/sections/real_space_formalism.tex
Edits to sec3.4 still has some holes and needs some refining

\bar O_{E/B} = 0.5 \bmat \mathcal{I}_{r} & \mathcal{I}_{i} \\  -\mathcal{I}_{i}  & \mathcal{I}_{r} \emat \pm \bmat \mathcal{D}_{r} & \mathcal{D}_{i} \\  \mathcal{D}_{i}  & - \mathcal{D}_{r} \emat \,,\\

@@

\mathcal{I} &=& \mathcal{I}_{r} + i \mathcal{I}_{i} = \sum_{\ell m} {_{-2}Y}_{\ell m}(\hat n_i) {_{-2}Y}^*_{\ell m}(\hat n_j) \,, \\
\mathcal{D}  &=& \mathcal{D}_{r} + i\mathcal{D}_{i} = \sum_{\ell m} {_2Y}_{\ell m}(\hat n_i) {_{-2}Y}^*_{\ell m}(\hat n_j) \,.

@@

\mathcal{I}(\hat{n}_i, \hat{n}_j) &=& \sum_{\ell} \sqrt{\frac{2\ell+1}{ 4 \pi}}{_{-2}Y}_{\ell2}(\beta_{ij}, \alpha_{ij}) ~ \rm{e}^{i2 \gamma_{ij}} \label{eq:healpix-compatible-i} = \mathcal{I}_r + i \mathcal{I}_i \,, \\
\mathcal{I}_r + i \mathcal{I}_i &=& \Big [ \cos(2 \alpha_{ij} +  2\gamma_{ij}) + i \sin(2 \alpha_{ij} +  2 \gamma_{ij}) \Big]   {{}_{\mi}f}(\beta_{ij},\ell_{\rm min},\ell_{\rm max}) \,,

@@

\mathcal{D}(\hat{n}_i, \hat{n}_j) &=& \sum_{\ell} \sqrt{\frac{2\ell+1}{ 4 \pi}}{_2Y}_{\ell 2}(\beta_{ij}, \alpha_{ij}) ~ \rm{e}^{- i2 \gamma_{ij}} \label{eq:healpix-compatible-m} =\mathcal{D}_r + i \mathcal{D}_i \,, \\
\mathcal{D}_r + i \mathcal{D}_i &=&  \Big [ \cos(2 \alpha_{ij} - 2\gamma_{ij}) + i \sin(2 \alpha_{ij} -  2 \gamma_{ij}) \Big]   {{}_{\md}f}(\beta_{ij},\ell_{\rm min},\ell_{\rm max}) \,,

@@

\bmat Q_i \\ U_i  \emat_{E/B} &=& \sum_{j=1}^{N_{\rm pix}} \Bigg\lbrace {{}_{\mi}f}(\beta_{ij},\ell_{\rm min},\ell_{\rm max}) \bmat \cos(2 \alpha_{ij} + 2\gamma_{ij}) & \sin(2\alpha_{ij} +2 \gamma_{ij}) \\  -\sin(2\alpha_{ij} +2 \gamma_{ij})  & \cos(2 \alpha_{ij} + 2 \gamma_{ij}) \emat  \bmat Q_j \\ U_j  \emat  \\ &\pm& {}_{\md}f(\beta_{ij},\ell_{\rm min},\ell_{\rm max}) \bmat \cos(2 \alpha_{ij} - 2\gamma_{ij}) &  \sin(2\alpha_{ij} - 2 \gamma_{ij}) \\  \sin(2\alpha_{ij} - 2 \gamma_{ij})  & - \cos(2 \alpha_{ij} - 2 \gamma_{ij}) \emat  \bmat Q_j \\ U_j  \emat \Bigg\rbrace 0.5 \Delta\Omega  \,, \nonumber 

@@

{}_{+2}X_{E/B} &=&0.5 \Delta \Omega\sum_{j=1}^{N_{\rm pix}}  {{}_{\mi}f}(\beta_{ij}) e^{-i2 (\alpha_{ij} + \gamma_{ij})} {}_{+2}X_j \pm {{}_{\md}f}(\beta_{ij}) e^{i2 (\alpha_{ij} - \gamma_{ij})} {}_{+2}X_j^*\,, \label{eq:qu2equbqu_explicit_convolution} \\

@@

where in \eq{eq:qu2equbqu_explicit_convolution} we have suppressed the explicit multipole dependence of functions $_{\pm 2}f$ for brevity and in \eq{eq:qu2equbqu_convolution} $\circ$ denotes a convolution. \comment{Here it will be nice to interpret $e^{i2\gamma}[E-iB]$. We don't understand this as of now.}

\comment{Recheck the math described below}
The function $\mathcal{I}$ is a band limited version of the delta function ($\lim_{\ell \rightarrow \infty} \mathcal{I} = \delta(\hat{n}_i - \hat{n}_j)$). When interpreted as a matrix it is a band limited version of the identity matrix. Though it has non vanishing off diagonal elements ($\mathcal{I} \neq 0 $ when $\hat{n}_i \neq \hat{n}_j$) owing to the band limit, for all practical purposes $\mathcal{I}$ acts like an identity operator as is confirmed by the following set of identities: (i) $\mathcal{I}*\mathcal{I}=\mathcal{I}$ ; (ii) $\mathcal{D}*\mathcal{I}=\mathcal{D}$. Also $\mathcal{D}^*$ is the inverse of $\mathcal{D}$ in this band limited sense: $\mathcal{D}^**\mathcal{D}=\mathcal{I}$. It is useful to note that the operator $\mathcal{D}$ is a complex but symmetric matrix and $\mathcal{I}$ is an Hermitian operator. Using these properties of the operators $\mathcal{I}$ and $\mathcal{D}$, one can verify that the real space operators satisfy the following identities,
%
\begin{subequations}
\beqry
\bar O_E * \bar O_E &=& \bar O_E ~~;~~ \bar O_B * \bar O_B = \bar O_B \,, \\
\bar O_E*\bar O_B &=& 0 \,,\\
\bar O_E + \bar O_B &=& \mathcal{I} \,,
\eeqry
\end{subequations}
%
which are the real space analogues of \eq{eq:har_op_prop}. While testing the above stated identities one encounters terms like $\mathcal{D}*\mathcal{I}^*,\mathcal{I}^**\mathcal{I} \textrm{ and }  \mathcal{I}*\mathcal{I}^*$ which cannot be simply interpreted, but they always occur in pairs with opposite signs, hence exactly canceling each other.

>>>>

\bar O_{E/B} = 0.5 \bmat \mathcal{I}_{r} & \mathcal{I}_{i} \\  -\mathcal{I}_{i}  & \mathcal{I}_{r} \emat \pm 0.5 \bmat \mathcal{D}_{r} & \mathcal{D}_{i} \\  \mathcal{D}_{i}  & - \mathcal{D}_{r} \emat \,,\\

@@

\mathcal{I} (\hat{n}_e,\hat{n}_q) &=& \mathcal{I}_{r} + i \mathcal{I}_{i} = \sum_{\ell m} {_{-2}Y}_{\ell m}(\hat n_e) {_{-2}Y}^*_{\ell m}(\hat n_q) \,, \\
\mathcal{D}(\hat{n}_e,\hat{n}_q)  &=& \mathcal{D}_{r} + i\mathcal{D}_{i} = \sum_{\ell m} {_2Y}_{\ell m}(\hat n_e) {_{-2}Y}^*_{\ell m}(\hat n_q) \,.

@@

\mathcal{I}(\hat{n}_e, \hat{n}_q) &=& \sum_{\ell} \sqrt{\frac{2\ell+1}{ 4 \pi}}{_{-2}Y}_{\ell2}(\beta_{qe}, \alpha_{qe}) ~ \rm{e}^{i2 \gamma_{qe}} \label{eq:healpix-compatible-i} = \mathcal{I}_r + i \mathcal{I}_i \,, \\
\mathcal{I}_r + i \mathcal{I}_i &=& \Big [ \cos(2 \alpha_{qe} +  2\gamma_{qe}) + i \sin(2 \alpha_{qe} +  2 \gamma_{qe}) \Big]   {{}_{\mi}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max}) \,,

@@

\mathcal{D}(\hat{n}_q, \hat{n}_e) &=& \sum_{\ell} \sqrt{\frac{2\ell+1}{ 4 \pi}}{_2Y}_{\ell 2}(\beta_{qe}, \alpha_{qe}) ~ \rm{e}^{- i2 \gamma_{qe}} \label{eq:healpix-compatible-m} =\mathcal{D}_r + i \mathcal{D}_i \,, \\
\mathcal{D}_r + i \mathcal{D}_i &=&  \Big [ \cos(2 \alpha_{qe} - 2\gamma_{qe}) + i \sin(2 \alpha_{qe} -  2 \gamma_{qe}) \Big]   {{}_{\md}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max}) \,,

@@

\bmat Q_e \\ U_e  \emat_{E/B} &=& \sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace {{}_{\mi}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max}) \bmat \cos(2 \alpha_{qe} + 2\gamma_{qe}) & \sin(2\alpha_{qe} +2 \gamma_{qe}) \\  -\sin(2\alpha_{qe} +2 \gamma_{qe})  & \cos(2 \alpha_{qe} + 2 \gamma_{qe}) \emat  \bmat Q_q \\ U_q  \emat  \\ &\pm& {}_{\md}f(\beta_{qe},\ell_{\rm min},\ell_{\rm max}) \bmat \cos(2 \alpha_{qe} - 2\gamma_{qe}) &  \sin(2\alpha_{qe} - 2 \gamma_{qe}) \\  \sin(2\alpha_{qe} - 2 \gamma_{qe})  & - \cos(2 \alpha_{qe} - 2 \gamma_{qe}) \emat  \bmat Q_q \\ U_q  \emat \Bigg\rbrace 0.5 \Delta\Omega  \,, \nonumber 

@@

{}_{+2}X_{E/B}(\hat{n}_e) &=&0.5 \Delta \Omega\sum_{q=1}^{N_{\rm pix}}  {{}_{\mi}f}(\beta_{qe}) e^{-i2 (\alpha_{qe} + \gamma_{qe})} {}_{+2}X(\hat{n}_q) \pm {{}_{\md}f}(\beta_{qe}) e^{i2 (\alpha_{qe} - \gamma_{qe})} {}_{+2}X(\hat{n}_q)^* \,, \nonumber \\

@@

where we have suppressed the explicit multipole dependence of functions ${}_{\mdi}f$ for brevity. \revisit{Comment on the radiating and convolution form of the operators.}

\comment{Recheck the math described below}
The operator $\mathcal{I}$ is a band limited version of the delta function ($\lim_{\ell \rightarrow \infty} \mathcal{I} = \delta(\hat{n}_i - \hat{n}_j)$) for spin-2 fields. When interpreted as a matrix it is a band limited version of the identity matrix. Though it has non-vanishing off diagonal elements ($\mathcal{I} \neq 0 $ when $\hat{n}_i \neq \hat{n}_j$) owing to the band limit, for all practical purposes $\mathcal{I}$ acts like an identity operator as is confirmed by the following set of identities: (i) $\mathcal{I}*\mathcal{I}=\mathcal{I}$ ; (ii) $\mathcal{D}*\mathcal{I}=\mathcal{D}$. Also $\mathcal{D}^*$ is the inverse of $\mathcal{D}$ in this band limited sense: $\mathcal{D}^**\mathcal{D}=\mathcal{I}$. It is useful to note that the operator $\mathcal{D}$ is a complex but symmetric matrix and $\mathcal{I}$ is an Hermitian operator. Using these properties of the operators $\mathcal{I}$ and $\mathcal{D}$, one can verify that the real space operators satisfy the following identities,
%
\begin{subequations}
\beqry
\bar O_E * \bar O_E &=& \bar O_E ~~;~~ \bar O_B * \bar O_B = \bar O_B \,, \\
\bar O_E*\bar O_B &=& 0 \,,\\
\bar O_E + \bar O_B &=& \mathcal{I} \,,
\eeqry
\end{subequations}
%
which are the real space analogues of \eq{eq:har_op_prop}. While testing the above stated identities one encounters terms like $\mathcal{D}*\mathcal{I}^*,\mathcal{I}^**\mathcal{I} \textrm{ and }  \mathcal{I}*\mathcal{I}^*$ which cannot be simply interpreted, but they always occur in pairs with opposite signs, hence exactly cancel each other.
